Person with advantage functionality for `hyper3` objects: `pwa3()`

### 2025-12-09 inst/pwa3.Rmd
pole position analysis

```{r maxp4}
H <- lam4(optimal_sol4$maximum, givelike=TRUE)
H
summary(H)
maxp(H)
```

Above we see that the evaluate is very high for `c` having a BT
strength of about 0.6.  This is consistent with our observations for
twice we see `c` winning [that is, coming first] without any help.

@@

>>>>

@@

```{r maxp4,cache=TRUE}
H <- lam4(optimal_sol4$maximum, givelike=TRUE)
H
summary(H)
maxp(H)
```

Above we see that the evaluate is very high for `c` having a BT
strength of about 0.6.  This is consistent with our observations for
twice we see `c` winning [that is, coming first] without any help.
Is there statistical evidence to support this assertion?

The simplest approach would be to consider the likelihood function at
$\hat{\lambda}\simeq 16.705$:

```{r stat4, cache=TRUE}
f <- function(o){samep.test(H, c(o, "c"))$statistic}
stat4 <- c(a=f("a"), b=f("b"), d=f("d"), e=f("e"))
```

```{r}
stat4
```

Above we see that there is no strong evidence for `c` being the
strongest, on the grounds that we are unable to reject $H_0\colon
p_b=p_c$ [the support difference is about 0.74, short of the two-units
of support criterion required for rejection], at least with
$\lambda=16.705$.

# Another further idea {-}

Now we have a model for the starting grid in formula 1.   In the
following function we specify the finishing order and also the
starting grid order.

```{r lam5}

tolam <- function(lam,o){setNames(lam^(seq_along(o)-1),o)}

lam5 <- function(lam, givelike=FALSE){
    race_help <- 
        race(c("a","b","d","c","e")) |> pwa3(tolam(lam, c("a","b","c","d","e"))) +
        race(c("a","c","b","d","e")) |> pwa3(tolam(lam, c("c","b","a","d","e"))) +
        race(c("a","d","c","b","e")) |> pwa3(tolam(lam, c("a","c","b","d","e"))) +
        race(c("e","c","b","a","d")) |> pwa3(tolam(lam, c("a","c","b","e","d"))) +
        race(c("b","c","e","a","d")) |> pwa3(tolam(lam, c("c","b","e","d","a"))) +
        race(c("a","d","c","b","e")) |> pwa3(tolam(lam, c("a","b","c","d","e")))

    if(givelike){
        return(race_help)
    } else {
        return(maxp(race_help, give=TRUE)$value)
    }
}
```

```{r lam5calculator, cache=TRUE}
l <- exp(seq(from= -2, to=log(1.5), len=15))
like5 <- sapply(l,lam5)
optimal_sol5 <- optimize(lam5, exp(c(-1,0)), maximum=TRUE)
```

```{r lamfminus5, cache=TRUE}
f <- function(l){lam5(l) - optimal_sol5$objective + 2}
lower <- uniroot(f, interval = c(exp(-1.5), exp(-1.0)))
upper <- uniroot(f, interval = c(exp(-0.5), exp(+0.5)))
```

```{r plotlamf5, echo=TRUE, fig.cap="Log-likelihood function for lambda, the pole position measure"}
plot(log(l), like5 - max(like5), type='b', pch=16)
abline(h = c(0, -2), col='gray')
abline(v = log(optimal_sol5$maximum))
abline(v = 0, lty=2)
segments(x0 = log(c(lower$root, upper$root)), y0 = -2.5, y1 = -1.5, col='red')
```
